---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?

AI, or Artificial Intelligence, involves creating computer systems that can perform tasks requiring human-like intelligence. There are two main types: Narrow AI, designed for specific tasks, and General AI, with broad cognitive abilities like humans (still a theoretical goal). AI systems can be categorized by their functions: Artificial Narrow Intelligence (ANI) for specific tasks, Artificial General Intelligence (AGI) for broader understanding, and Artificial Superintelligence (ASI) as a theoretical superintelligent state. AI is applied in various fields using technologies like machine learning and deep learning, raising ongoing ethical and research challenges.

2. What is ChatGPT?

ChatGPT is a language model developed by OpenAI. It is part of the GPT (Generative Pre-trained Transformer) family, specifically based on the GPT-3.5 architecture. ChatGPT is designed for generating human-like text responses based on the input it receives. It can be used for a variety of natural language processing tasks, including answering questions, providing information, generating creative content, and engaging in conversation.

One notable feature of ChatGPT is its ability to understand and generate contextually relevant responses in a conversational style. It has been trained on a diverse range of internet text and is capable of producing coherent and contextually appropriate replies.

It's important to note that while ChatGPT demonstrates impressive language generation capabilities, it may not always provide accurate or up-to-date information, and its responses are based on patterns learned during training. Users should exercise caution and critical judgment when interacting with AI models like ChatGPT.

3. What is Github Copilot?

GitHub Copilot is an AI-driven code completion tool developed by GitHub and OpenAI. It assists developers by suggesting and autocompleting code as they write, using the OpenAI Codex model. The tool supports multiple programming languages and aims to understand the developer's context for more relevant suggestions, potentially speeding up the coding process. Developers should review suggestions to ensure code correctness and adherence to standards.

4. What is the rarest color in nature?

The perception of rarity in colors can be subjective, and it depends on various factors such as context and interpretation. However, one color often considered rare in nature is "blue." True blue pigments are less common in living organisms compared to other colors. While there are certainly examples of blue in nature, such as certain flowers, birds, and butterflies, the true production of blue pigments can be a more complex biochemical process, making it relatively less common than other colors like green or brown. Some notable examples of naturally occurring blue pigments include the wings of the blue morpho butterfly and certain species of birds.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("data/netflix_titles.csv")

```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
glimpse(netflix)

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- netflix %>%
  filter(type=="Movie")
```


Which director has the most movies on netflix?
```{r}
count(netflix_movies, director, sort = T)

```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

[insert your prompt here]
```{r}
filter(count(netflix_movies, director, sort = T), !is.na(director))
```



Which director had the most movies by release year and remove the NA's?
```{r}

```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


I want to count the number of movies made by each director in a movie dataset; however, movies directed by multiple directors are counted separately from the movies each director made individually. How would I fix this? Here is the code I used: "filter(count(netflix_movies, director, sort = T), !is.na(director))"
```{r}
result <- netflix_movies %>%
  filter(!is.na(director)) %>%
  separate_rows(director, sep = ", ") %>%
  group_by(director) %>%
  summarize(num_movies = n(), .groups = 'drop') %>%
  arrange(desc(num_movies))
```



Make a bar plot of the top directors with the most movies on netflix
```{r}
# Assuming 'netflix_movies' is your dataset
top_directors <- netflix_movies %>%
  filter(!is.na(director)) %>%
  group_by(director) %>%
  summarize(num_movies = n(), .groups = 'drop') %>%
  arrange(desc(num_movies)) %>%
  head(10)  # Adjust the number to show the top N directors

# Create a bar plot
ggplot(top_directors, aes(x = reorder(director, -num_movies), y = num_movies)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Top Directors with the Most Movies on Netflix",
       x = "Director",
       y = "Number of Movies") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*

[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt

```{r}
# Install required packages if not installed
if (!requireNamespace("shiny", quietly = TRUE)) {
  install.packages("shiny")
}

if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

library(shiny)
library(ggplot2)

# Assuming 'netflix_movies' is your dataset
# Replace this with your actual dataset
# Sample data for illustration
set.seed(123)
netflix_movies <- data.frame(
  director = sample(c("DirectorA", "DirectorB", "DirectorC"), 100, replace = TRUE),
  num_movies = sample(1:10, 100, replace = TRUE)
)

# Shiny UI
ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  sidebarLayout(
    sidebarPanel(),
    mainPanel(
      plotOutput("barplot")
    )
  )
)

# Shiny Server
server <- function(input, output) {
  # Function to get the top directors
  getTopDirectors <- function(dataset, top_n = 10) {
    dataset %>%
      filter(!is.na(director)) %>%
      group_by(director) %>%
      summarize(num_movies = sum(num_movies), .groups = 'drop') %>%
      arrange(desc(num_movies)) %>%
      head(top_n)
  }

  # Reactive function to update the plot
  output$barplot <- renderPlot({
    top_directors <- getTopDirectors(netflix_movies)
    
    # Create a bar plot
    ggplot(top_directors, aes(x = reorder(director, -num_movies), y = num_movies)) +
      geom_bar(stat = "identity", fill = "skyblue", color = "black") +
      labs(title = "Top Directors with the Most Movies on Netflix",
           x = "Director",
           y = "Number of Movies") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  })
}

# Run the Shiny App
shinyApp(ui = ui, server = server)

```


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
