---
title: "Lab 4 Homework"
author: "Jaskaran Halait"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions (complete questions 1 - 6)
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  
Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

**1. Load the data into a new object called `homerange`.**
```{r}
homerange <- read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```

**2. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.**  
```{r}
dim(homerange)
```
The data frame contains 569 observations and 24 variables.

```{r}
names(homerange)
```
The column names are as follows: taxon, common.name, class, order, family, genus, species, primarymethod, N, mean.mass.g, log10.mass, alternative.mass.reference, mean.hra.m2, log10.hra, hra.reference, realm, thermoregulation, locomotion, trophic.guild, dimension, preymass, log10.preymass, PPMR, and prey.size.reference.

```{r}
glimpse(homerange)
```
The variables taxon, common.name, class, order, family, genus, species, primarymethod, N, alternative.mass.reference, hra.reference, realm, thermoregulation, locomotion, trophic.guild, and prey.size.reference are all the character type. All other variables are double precisions (numerics).

```{r}
summary(homerange)
```

**3. Change the class of the variables `taxon` and `order` to factors and display their levels.**  
```{r}
#Changing the variables 'taxon' and 'order' from the character type to factors:
homerange$taxon <- as.factor(homerange$taxon)
homerange$order <- as.factor(homerange$order)
#Double-checking that the commands were carried out as intended:
homerange
```

```{r}
#Listing the levels of the 'taxon' variable:
levels(homerange$taxon)
```

```{r}
#Listing the levels of the 'order' variable:
levels(homerange$order)
```

**4. What taxa are represented in the `homerange` data frame? Make a new data frame `taxa` that is restricted to taxon, common name, class, order, family, genus, species.**  
```{r}
taxa <- select(homerange, "taxon", "common.name", "class", "order", "family", "genus", "species")
taxa
```

**5. The variable `taxon` identifies the common name groups of the species represented in `homerange`. Make a table the shows the counts for each of these `taxon`.**  
```{r}
table(homerange$taxon)
```

**6. The species in `homerange` are also classified into trophic guilds. How many species are represented in each trophic guild.**  
```{r}
table(homerange$trophic.guild)
```
342 of the 569 species in the homerange data set can be classified as carnivores, while the remaining 227 are herbivores.

**7. Make two new data frames, one which is restricted to carnivores and another that is restricted to herbivores.**  
```{r}
#Create the carnivore data frame:
carnivore_subset <- filter(homerange, trophic.guild=="carnivore")
```

```{r}
#Create the herbivore data frame:
herbivore_subset <- filter(homerange, trophic.guild=="herbivore")
```


**8. Do herbivores or carnivores have, on average, a larger `mean.hra.m2`? Remove any NAs from the data.**  
```{r}
#Calculate the carnivore average mean.hra.m2:
mean(carnivore_subset$"mean.hra.m2", na.rm = T)
```

```{r}
#Calculate the herbivore average mean.hra.m2:
mean(herbivore_subset$"mean.hra.m2", na.rm = T)
```

**9. Make a new dataframe `owls` that is limited to the mean mass, log10 mass, family, genus, and species of owls in the database. Which is the smallest owl? What is its common name? Do a little bit of searching online to see what you can learn about this species and provide a link below** 
```{r}
#Generating the data frame of interest:
owls <- filter(homerange, order=="strigiformes")
owls_subset <- select(owls, "common.name", "mean.mass.g", "log10.mass", "family", "genus", "species")

#Finding the smallest owl
?min()
min(owls_subset$mean.mass.g)
filter(owls_subset, mean.mass.g==61.32)
```

The smallest of the owls by average mass is the Eurasian pygmy owl. [Here is a link to its wikipedia entry

**10. As measured by the data, which bird species has the largest homerange? Show all of your work, please. Look this species up online and tell me about it!**.  
```{r}

```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   