---
title: "Lab 3 Homework"
author: "Jaskaran (Jaski) Halait"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

### Mammals Sleep  
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R. The name of the data is `msleep`.  
```{r}
?msleep
```

This data set was taken from the journal article "A quantitative, theoretical framework for understanding mammalian sleep" by V. M. Savage and G. B. West.

2. Store these data into a new data frame `sleep`.  
```{r}
sleep <- msleep
sleep
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
str(sleep)
```

The data frame is 83 rows (observations) by 11 columns (variables). This was determined using the "str()" function.

4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
is.na(sleep)
```

Yes, there are missing values in the data frame, according to the function "is.na()".

5. Show a list of the column names is this data frame.
```{r}
?names
names(sleep)
```

6. How many herbivores are represented in the data?  
```{r}
herbivores <- filter(sleep, vore=="herbi")
herbivores
```

32 mammals in the data set are herbivores.

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 19kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
#Creating the small mammal data frame
small <- filter(sleep, bodywt<=19)
small

#Creating the large mammal data frame
large <- filter(sleep, bodywt>=200)
large
```

8. What is the mean weight for both the small and large mammals?
```{r}
#Mean weight of small mammals
average_weight_small <- mean(small$bodywt)
average_weight_small
```

The mean weight of the small mammals is 1.797847 kg. 

```{r}
#Mean weight of large mammals
average_weight_large <- mean(large$bodywt)
average_weight_large
```

The mean weight of the large mammals is 1747.071 kg. 

9. Using a similar approach as above, do large or small animals sleep longer on average?  

```{r}
#Mean total sleep for small mammals
average_sleep_small <- mean(small$sleep_total)
average_sleep_small
```

```{r}
#Mean total sleep for large mammals
average_sleep_large <- mean(large$sleep_total)
average_sleep_large
```

On average, small mammals sleep much longer (11.79 hours) than large mammals do (3.3 hours).

10. Which animal is the sleepiest among the entire data frame?

```{r}
#Find the highest time spent sleeping by any mammal in the data frame
?max
max_sleep <- max(sleep$sleep_total)
max_sleep
```

```{r}
#Find the mammal that sleeps this long
sleepiest_mammal <- filter(sleep, sleep_total==19.9)
sleepiest_mammal
sleepiest_mammal[1,1]
```
The little brown bat sleeps the longest of all the mammals in the data frame.

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   