---
title: "lab_06_warmup"
output: 
  html_document: 
    keep_md: true
date: "2024-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lab 06 Warmup
#### Load any required packages
```{r}
library("tidyverse")
```
#### 1 Load the 'bison.csv' data.
```{r}
bison <- read_csv("data/bison.csv")
```

#### 2 What are the dimesions and structure of the data?
```{r}
dim(bison)
str(bison)
```

#### 3 We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.

```{r}
bison_subset <- select(bison, "data_code", "animal_code", "animal_sex", "animal_weight", "animal_yob")
```


#### 4 Pull out the animals born between 1980-1990.
```{r}
bison_1980_to_1990 <- filter(bison_subset, between(animal_yob, 1980, 1990))
```

#### 5 How many male and female bison are represented between 1980-1990?
```{r}
table(bison_1980_to_1990$animal_sex)
```

#### 6 Between 1980-1990, were males or females larger on average? Males
```{r}
males <- filter(bison_1980_to_1990, animal_sex=="M")
  mean(males$animal_weight, na.rm = T)
  
females <- filter(bison_1980_to_1990, animal_sex=="F")
  mean(females$animal_weight, na.rm = T)
```

